---
title: "model_results"
author: "Scott Klasek"
date: "`r Sys.Date()`"
output: github_document
---
## Purpose
Summarize model information here because the document 45_SEM_stuff is too big and scary.  

## Setup
#### load libraries
```{r}
packages <- c("tidyverse", "phyloseq", "speedyseq", "patchwork", "compositions", "piecewiseSEM", "lme4")
invisible(lapply(packages, require, character.only = TRUE))
```

#### load data
```{r}
# lists of phyloseq objects
# these include metadata updated in doc 39.Rmd
# they are ALL years, ASVs and samples NOT subsetted or transformed
its.ps.list <- readRDS(file = "/Users/klas0061/Desktop/UMN/treatment_variance_modeling/its.ps.list")
its.ps.list <- its.ps.list[c(1:7,9)] # remove Larkin samples
bact.ps.list <- readRDS(file = "/Users/klas0061/Desktop/UMN/treatment_variance_modeling/bact.ps.list")
bact.ps.list <- bact.ps.list[c(1:7,9)] # remove Larkin samples

# jim's spreadsheet of soil chemical metadata
jim.info <- read.csv(file="/Users/klas0061/Desktop/UMN/jim_info/PSHP_ALL_obj_1_all_data_2024_02_29.csv") # import
```

#### reformat soil chemical metadata so it can be added to phyloseq objects using merge_with_jim_data       
```{r}
# basic clean-up stuff
colnames(jim.info) <- jim.info[1,] # fix column names (there are two header cols)
jim.info <- jim.info[2:nrow(jim.info),] # omit redundant column name 

for (i in c(2:3,5:6)){jim.info[,i] <- as.numeric(jim.info[,i])} # convert columns that should be numeric into numeric

# remove extraneous year columns that correspond to yield data only
jim.info <- jim.info[,c(1:101,103:129,131:ncol(jim.info))]

# select data columns to merge by, and to add (an example, can always select more later)
jim.info.s <- jim.info %>% dplyr::select(State, Objective, Rotation, Plot, Year, Month, 
                                  pH, `Nitrate-N (ppm)`, `NH4-N (ppm)`, `P-Olsen (ppm)`, `P-Bray (ppm)`,
                                  `OM (%)`, `Solvita (ppm)`, `Total C (%)`, `Total organic C (%)`, `POX-C (ppm)`) 

for (i in 7:16){jim.info.s[,i] <- as.numeric(jim.info.s[,i])} # again, convert columns that should be numeric into numeric

# change OR Spring 2020 month to 4 so Cultivar (whether field is in potato or not) is not NA
jim.info.s[which(jim.info.s$State=="OR" & jim.info.s$Year==20 & jim.info.s$Month=="3"),"Month"] <- "4" 

jim.info.s$Month <- as.numeric(jim.info.s$Month) # more character to numeric
```

#### define functions
```{r}
# drop_ghost_asvs drops ASVs that are not present in any samples- but are still kept because you've 
# subsetted samples within a larger phyloseq object. This returns the phyloseq object with the subsetted count table.
drop_ghost_asvs <- function(ps){
  
  # extract the count table from the phyloseq, with ASVs as columns
    if(taxa_are_rows(ps)){counts <- t(ps@otu_table)} else{ # write count table with ASVs as rows
    counts <- ps@otu_table
    }
  # remove empty rows corresponding to ASVs that are present in 0 samples
  counts <- counts[,colSums(counts)>0]
  
  # write the count table back in
  otu_table(ps) <- counts # drops all zero-count ASVs from the tax table and refseq as well, in contrast to 'ps@otu_table <- counts'
  return(ps)
}

# subset_occupancy drops ASVs detected in less than a certain proportion of the samples
# all it takes is a phyloseq object. it drops the ASVs below the occupancy cutoff, and moves their counts to 
# a "summing" column. Tax table now has a summing column as well. Output is a phyloseq object. 
subset_occupancy <- function(ps, occ_cutoff){
  
  # first, calculate occupancy of each ASV (proportion of samples detected in, from 0 to 1)
  occ <- vector("numeric")
    for (i in 1:ncol(ps@otu_table)) {
      occ[i] <- sum(ps@otu_table[,i] != 0)/nrow(ps@otu_table[,i])
    }
  # write ASV names
  names(occ) <- colnames(ps@otu_table)
    
  # select ASVs to keep based on occupancy 
  keep.asvs <- occ[which(occ >= (occ_cutoff))] # keep only the ASVs above the cutoff
  
  # print some useful information
  cat((length(keep.asvs)), "features are kept. \n")
  
  # subset the count table to maintain compositionality
  keep.table <- ps@otu_table[,names(keep.asvs)] # the kept ASVs, which does not include a summing column
  drop.table <- ps@otu_table[,setdiff(colnames(ps@otu_table), names(keep.asvs))] # the dropped ASVs
  summing <- rowSums(drop.table) # add up counts for dropped features to make a summing column
  keep.new <- cbind(keep.table, summing) # add new summing to the keep table
  
  # make a subsetted tax_table
  taxa <- tax_table(ps)[names(keep.asvs),] # with only the kept ASVs over the occupancy cutoff
  summing <- c("Summing", rep(NA, times = ncol(taxa)-1))
  taxa2 <- rbind(taxa, summing)
    
  # put the new count table and tax table into a new ps object
  # we must say goodbye to our refseq here because "summing" doesn't have a sequence associated with it.
  new.ps <- phyloseq(sample_data(ps),
                     otu_table(keep.new, taxa_are_rows = FALSE),
                     tax_table(taxa2))
  return(new.ps)
}

# transform_clr takes a phyloseq object and transforms the count table using a centered-log-ratio
# it spits out a phyloseq object with transformed counts.
transform_clr <- function(ps){
  counts.clr <- as.matrix(clr(ps@otu_table))
  ps.new <- phyloseq(otu_table(counts.clr, taxa_are_rows = FALSE),
                     tax_table(ps),
                     sample_data(ps))
  cat("Counts transformed with CLR. \n")
  return(ps.new)
}

# merge_with_jim_data takes as input a phyloseq object and a dataframe of objective 1 data that Jim has curated, which can be subsetted 
# it writes data from Jim's spreadsheet into the corresponding phyloseq object, based on combinations
# make sure jim's data is cleaned up according to the steps above!
merge_with_jim_data <- function(ps, jim.info.subset){
  df <- left_join(data.frame(ps@sam_data), jim.info.subset, 
                by = c("state" = "State", "objective" = "Objective", "rotation" = "Rotation",
                       "plot" = "Plot", "year" = "Year", "month" = "Month")) # merges 
  #the unique combo of columns from phyloseq sample data and jim's data
  rownames(df) <- rownames(ps@sam_data) # ok as long as the rows are in the same order, which they are
  
  # put df into the sample_data.
  sample_data(ps) <- df # this also automatically drops the ghost ASVs and refsequences from the ps object
  return(ps)
}
```


## Microbiome-informed models   

### Scenario 1: OR 2-yr bacterial consortia in response to amendment    
From the 2-yr rotation only  

```{r}
# begin by subsetting the phyloseq object by site, rotation, and year
ps6 <- bact.ps.list[["OR.16S.ps"]] %>% 
  subset_samples(year == 22 & rotation == 2 & !is.na(Total.yield) & general_category != "Fumigated") 

# subset soil chemical data, and transfer spring measurements of OM % to their corresponding summer samples 
sub.df <- jim.info.s %>% filter(State == "OR" & Year == 22)
sub.df <- sub.df %>% group_by(Plot) %>% fill(`OM (%)`)

# import this soil data into the phyloseq object
ps6 <- merge_with_jim_data(ps6, sub.df)

# define ASVs to model
or.amd.bact.summer.asvs <- c("ASV1254","ASV21630", "ASV2399",
                          "ASV3632",  "ASV4948", "ASV5076",  "ASV7919", "ASV792")

# keep only the summer samples
ps6 <- ps6 %>% subset_samples(season == "Summer") %>% 
  drop_ghost_asvs() %>%
  subset_occupancy(0.5) %>% 
  transform_clr() 

# add in the abundance of the 2 yr rotation SUMMER associated ASVs to the sample data 
df1 <- cbind(data.frame(ps6@sam_data), ps6@otu_table[,or.amd.bact.summer.asvs])

# make a consortium column of the sum of the ASVs
df1$consortium <- rowSums(df1[,startsWith(colnames(df1), "ASV")])

### tidying up the dataframe 
# remove NA cols
df1 <- df1 %>% dplyr::select(-cover_2020, -green_cov_pct_pre_kill, -Total.C...., -Total.organic.C....)

# there was one outlier OM value at 2.8, next highest OM recorded in Oregon any time was 1.8
# also keep only Norkotah cultivar
df1.no <- df1 %>% filter(`OM....` < 2 & cultivar == "Norkotah")

# lm where amendment directly impacts yield as well (this cannot be modeled in SEM)
lm4 <- lm(Total.yield ~ Amended + consortium, df1.no)
summary(lm4)
extractAIC(lm4) # gives AIC of 87, which is lower than the PSEM

# calculate 
est_amd <- summary(lm(consortium ~ Amended, df1.no))$coefficients[2, 1]
est_cons <- summary(lm(Total.yield ~ consortium, df1.no))$coefficients[2, 1]

# Yield increase as an effect of treatment on microbiome (cwt/ac) is the product of the individual coefficients 
yield_inc <- est_amd * est_cons

# percent increase in total yield from treatment-related microbiome changes, relative to control  
pct_yield_inc <- est_amd * est_cons / mean(df1.no %>% filter(general_category == "Control") %>% pull(Total.yield)) * 100

# best model
summary(lm(Total.yield ~ consortium + Amended, df1.no))

# extract p- and R2 values
overall_p <- function(my_model) {
    f <- summary(my_model)$fstatistic
    p <- pf(f[1],f[2],f[3],lower.tail=F)
    attributes(p) <- NULL
    return(p)
}

pval <- overall_p(lm(Total.yield ~ consortium + Amended, df1.no))
rsq <- summary(lm(Total.yield ~ consortium + Amended, df1.no))[[8]]
```

### Scenario 2: ID 3-yr bacterial consortia in response to amendment    
From the 3-yr rotation only 

```{r}
# begin by subsetting the phyloseq object by site, rotation, and year
ps7 <- bact.ps.list[["ID.16S.ps"]] %>% 
  subset_samples(year == 22 & rotation == 3 & !is.na(Total.yield) & Fumigated == FALSE) 

# subset soil chemical data, and transfer spring measurements of OM % to their corresponding summer samples 
sub.df <- jim.info.s %>% filter(State == "ID" & Year == 22 & Rotation == 3)
sub.df <- sub.df %>% group_by(Plot) %>% fill(`OM (%)`)

# import this soil data into the phyloseq object
ps7 <- merge_with_jim_data(ps7, sub.df)

# define ASVs
id.amd.bact.summer.asvs <- c("ASV3781", "ASV12474", "ASV17047", "ASV2340",  "ASV2345",  
                             "ASV2589", "ASV4416",  "ASV4825",  "ASV4830")

# keep only the summer samples, remove any below 50% occupancy, and transform abundances
ps7 <- ps7 %>% subset_samples(season == "Summer") %>% 
  drop_ghost_asvs() %>%
  subset_occupancy(0.5) %>% 
  transform_clr() 

# add in the abundance of the summer associated ASVs to the sample data 
df2 <- cbind(data.frame(ps7@sam_data), ps7@otu_table[,id.amd.bact.summer.asvs])

# make a consortium column of the sum of the ASVs
df2$consortium <- rowSums(df2[,startsWith(colnames(df2), "ASV")])

# remember that Green Manure treatments are now classified as controls
df2[which(df2$amendment == "Green Manure"),"general_category"] <- "Control"

# control for cultivar: keep only Norkotahs
df2 <- df2 %>% filter(cultivar == "Norkotah")

# drop NA cols
df2 <- df2[, colSums(is.na(df2)) != nrow(df2)]

# convert amended treatment to 0/1 (NECESSARY FOR EXTRACTING MODEL INFO FROM PSEMs)
df2$Amended <- as.integer(as.logical(df2$Amended))

# SEM 2: amendment increases OM, OM and amendment increase consortium, and consortium increases total yields
psem2 <- psem(lm(`OM....` ~ Amended, df2),
               lm(consortium ~ `OM....` + Amended, df2),
               lm(Total.yield ~ consortium, df2))  

fisherC(psem2) # valid
coefs(psem2)

# extract p and R2 values 
psemrsq <- rsquared(psem2)[which(rsquared(psem2)$Response == "Total.yield"),"R.squared"]
rsq <- append(rsq, psemrsq)
pval <- append(pval, fisherC(psem2, conserve = T)[1,3])

# calculate 
est_amd <- summary(lm(consortium ~ Amended, df2))$coefficients[2, 1]
est_cons <- summary(lm(Total.yield ~ consortium, df2))$coefficients[2, 1]

# Yield increase as an effect of treatment on microbiome (cwt/ac) is the product of the individual coefficients 
yield_inc <- append(yield_inc, (est_amd * est_cons))

# percent increase in total yield from treatment-related microbiome changes, relative to control  
pct <- est_amd * est_cons / mean(df2 %>% filter(general_category == "Control") %>% pull(Total.yield)) * 100
pct_yield_inc <- append(pct_yield_inc, pct)
```

### Scenario 3: ME 2-yr ITS in response to amendment    
The 3-yr rotation is the one that had the yield increase, but the 2-yr one had the two amendment-influenced, yield-positive fungal ASVs.  

```{r}
# begin by subsetting the phyloseq object by site, rotation, and year
ps8 <- its.ps.list[["ME.ITS.ps"]] %>% 
  subset_samples(year == 22 & rotation == 2 & !is.na(Total.yield) & general_category != "Fumigated") 

# subset soil chemical data, and transfer spring measurements of OM % to their corresponding summer samples
sub.df <- jim.info.s %>% filter(State == "ME" & Year == 22 & Rotation == 2)
sub.df <- sub.df %>% group_by(Plot) %>% fill(`OM (%)`)

# import this soil data into the phyloseq object
ps8 <- merge_with_jim_data(ps8, sub.df)

# define ASVs
me.amd.its.spring.asvs <- c("ASV501", "ASV1906")

# keep only the spring samples, drop, and transform
ps8 <- ps8 %>% subset_samples(season == "Spring") %>% 
  drop_ghost_asvs() %>%
  subset_occupancy(0.5) %>% 
  transform_clr() 

# add in the abundance of the spring associated ASVs to the sample data 
df3 <- cbind(data.frame(ps8@sam_data), ps8@otu_table[,me.amd.its.spring.asvs])

# make a consortium column of the sum of the ASVs
df3$consortium <- rowSums(df3[,startsWith(colnames(df3), "ASV")])

# since we have two cultivars of nearly equal abundance, control for cultivar as a factor
df3$cultivar <- ifelse(df3$cultivar == "Burbank", 0, 
                       ifelse(df3$cultivar == "Caribou", 1, NA))

# model
summary(lm(Total.yield ~ Amended + consortium + cultivar, df3))

pval <- append(pval, overall_p(lm(Total.yield ~ Amended + consortium + cultivar, df3)))
rsq <- append(rsq, summary(lm(Total.yield ~ Amended + consortium + cultivar, df3))[[8]])

# calculate coefficients
# cultivar left out of consortium -> yield model because insignificant
est_amd <- summary(lm(consortium ~ Amended + cultivar, df3))$coefficients[2, 1]
est_cons <- summary(lm(Total.yield ~ consortium, df3))$coefficients[2, 1]

# Yield increase as an effect of treatment on microbiome (cwt/ac) is the product of the individual coefficients 
yield_inc <- append(yield_inc, (est_amd * est_cons))

# percent increase in total yield from treatment-related microbiome changes, relative to control  
pct <- est_amd * est_cons / mean(df3 %>% filter(general_category == "Control") %>% pull(Total.yield)) * 100
pct_yield_inc <- append(pct_yield_inc, pct)
```


### Scenario 4: MI 3-yr ITS in response to amendment   
```{r}
# begin by subsetting the phyloseq object by site, rotation, and year
ps9 <- its.ps.list[["MI.ITS.ps"]] %>% 
  subset_samples(year == 22 & rotation == 3 & !is.na(Total.yield) & general_category != "Control") 

# subset soil chemical data, and transfer spring measurements of OM % to their corresponding summer samples
sub.df <- jim.info.s %>% filter(State == "MI" & Year == 22 & Rotation == 3)
sub.df <- sub.df %>% group_by(Plot) %>% fill(`OM (%)`)

# import this soil data into the phyloseq object
ps9 <- merge_with_jim_data(ps9, sub.df)

# define ASVs to model
mi.amd.its.summer.asvs <- c("ASV1099", "ASV1194")

# keep only the spring samples, drop, and transform
ps9 <- ps9 %>% subset_samples(season == "Summer") %>% 
  drop_ghost_asvs() %>%
  subset_occupancy(0.5) %>% 
  transform_clr() 

# add in the abundance of the spring associated ASVs to the sample data 
df4 <- cbind(data.frame(ps9@sam_data), ps9@otu_table[,mi.amd.its.summer.asvs])

# make a consortium column of the sum of the ASVs
df4$consortium <- rowSums(df4[,startsWith(colnames(df4), "ASV")])

# convert amended treatment to 0/1
df4$Amended <- as.integer(as.logical(df4$Amended))

### tidying up the dataframe 
# remove NA cols
df4 <- df4 %>% dplyr::select(-cover_2019, -cover_2020, -green_cov_pct_pre_kill, -Total.C...., 
                             -Total.organic.C...., -VPPG, -Root.Lesion.100.cc, -rotation_comparison)

# encode cultivar as a dummy variable
df4$cultivar <- ifelse(df4$cultivar == "Superior", 0,
                       ifelse(df4$cultivar == "Burbank", 1, 2))

# SEM: amendment directly stimulates consortium, which increases yields
psem1 <- psem(lm(consortium ~ Amended + cultivar, df4),
              lm(Total.yield ~ consortium + cultivar, df4)) 

fisherC(psem1) # valid
coefs(psem1)

# extract p and R2 values
psemrsq <- rsquared(psem1)[which(rsquared(psem1)$Response == "Total.yield"),"R.squared"]
rsq <- append(rsq, psemrsq)
pval <- append(pval, fisherC(psem1, conserve = T)[1,3])

# calculate coefficients
# cultivar left out of consortium -> yield model because insignificant
est_amd <- summary(lm(consortium ~ Amended + cultivar, df4))$coefficients[2, 1]
est_cons <- summary(lm(Total.yield ~ consortium, df4))$coefficients[2, 1]

# percent increase in total yield from treatment-related microbiome changes, relative to control  
pct <- est_amd * est_cons / mean(df4 %>% filter(general_category == "Fumigated") %>% pull(Total.yield)) * 100

# Yield increase as an effect of treatment on microbiome (cwt/ac) is the product of the individual coefficients 
yield_inc <- append(yield_inc, (est_amd * est_cons))

# percent increase in total yield from treatment-related microbiome changes, relative to control  
pct_yield_inc <- append(pct_yield_inc, pct)
```

### Scenario 5: OR 3-yr ITS in response to fumigation   
```{r}
# begin by subsetting the phyloseq object by site, rotation, and year
ps1 <- its.ps.list[["OR.ITS.ps"]] %>% 
  subset_samples(year == 22 & rotation == 3 & !is.na(Total.yield) & general_category != "Amended") 

# subset soil chemical data, and transfer spring measurements of OM % to their corresponding summer samples 
sub.df <- jim.info.s %>% filter(State == "OR" & Year == 22)
sub.df <- sub.df %>% group_by(Plot) %>% fill(`OM (%)`)

# import this soil data into the phyloseq object
ps1 <- merge_with_jim_data(ps1, sub.df)

# keep only the spring samples, drop, and transform
ps1 <- ps1 %>% subset_samples(season == "Summer") %>% 
  drop_ghost_asvs() %>%
  subset_occupancy(0.5) %>% 
  transform_clr() 

# only one eukaryotic ASV associated with fumigation/yield, and it is 60 DAP
df5 <- cbind(data.frame(ps1@sam_data), ps1@otu_table[,"ASV2495"])

# encode cultivar as a dummy variable
df5$cultivar <- ifelse(df5$cultivar == "Burbank", 0,
                       ifelse(df5$cultivar == "Norkotah", 1, 2))

# Total.yield ~ 60 DAP abundance of ASV2495 + Fumigated + cultivar. Cultivars coded as dummy variables.
summary(lm(Total.yield ~ ASV2495 + Fumigated + cultivar, df5))

pval <- append(pval, overall_p(lm(Total.yield ~ ASV2495 + Fumigated + cultivar, df5)))
rsq <- append(rsq, summary(lm(Total.yield ~ ASV2495 + Fumigated + cultivar, df5))[[8]])

# calculate coefficients
# cultivar left out of consortium -> yield model because insignificant
est_amd <- summary(lm(ASV2495 ~ Amended + cultivar, df5))$coefficients[2, 1]
est_cons <- summary(lm(Total.yield ~ ASV2495, df5))$coefficients[2, 1]

# Yield increase as an effect of treatment on microbiome (cwt/ac) is the product of the individual coefficients 
yield_inc <- append(yield_inc, (est_amd * est_cons))

# percent increase in total yield from treatment-related microbiome changes, relative to control  
pct <- est_amd * est_cons / mean(df5 %>% filter(general_category == "Control") %>% pull(Total.yield)) * 100
pct_yield_inc <- append(pct_yield_inc, pct)
```

### Scenario 6: OR  3-yr bacteria in response to mustard incorporation  
```{r}
# begin by subsetting the phyloseq object by site, rotation, and year
ps <- bact.ps.list[["OR.16S.ps"]] %>% 
  subset_samples(year == 22 & rotation == 3 & !is.na(Total.yield) & general_category != "Fumigated") 

# subset soil chemical data, and transfer spring measurements of OM % to their corresponding summer samples
sub.df <- jim.info.s %>% filter(State == "OR" & Year == 22 & Rotation == 3)
sub.df <- sub.df %>% group_by(Plot) %>% fill(`OM (%)`)

# import this soil data into the phyloseq object
ps <- merge_with_jim_data(ps, sub.df)

# they're both in spring
or.must.bact.spring.asvs <- c("ASV270", "ASV2762", "ASV3009", "ASV3745", "ASV574")

# keep only the spring samples, drop, and transform
ps <- ps %>% subset_samples(season == "Spring") %>% 
  drop_ghost_asvs() %>%
  subset_occupancy(0.5) %>% 
  transform_clr() 

# add in the abundance of the spring associated ASVs to the sample data 
df6 <- cbind(data.frame(ps@sam_data), ps@otu_table[,or.must.bact.spring.asvs])

# make a consortium column of the sum of the ASVs
df6$consortium <- rowSums(df6[,startsWith(colnames(df6), "ASV")])

# drop non Norkotah cultivars
df6.no <- df6 %>% filter(cultivar == "Norkotah")

# drop NA cols
df6.no <- df6.no[, colSums(is.na(df6.no)) != nrow(df6.no)]

# convert Brassica to integer
df6.no$Brassica <- as.integer(as.logical(df6.no$Brassica))

# SEM: mustard increases nitrate, decreases pH, and increases cumulative target ASV abundances at planting, 
# abundances decrease yields. Best SEM included pH, but not nitrate. Omitted non-Norkotah cultivars.
psem3 <- psem(lm(pH ~ Brassica, df6.no),
              lm(consortium ~ Brassica, df6.no),
              lm(Total.yield ~ consortium + pH, df6.no))

fisherC(psem3) # valid
coefs(psem3)

# extract p and R2 values
psemrsq <- rsquared(psem3)[which(rsquared(psem3)$Response == "Total.yield"),"R.squared"]
rsq <- append(rsq, psemrsq)
pval <- append(pval, fisherC(psem3, conserve = T)[1,3])

# calculate coefficients
# cultivar left out of consortium -> yield model because insignificant
est_amd <- summary(lm(consortium ~ Brassica + pH, df6.no))$coefficients[2, 1]
est_cons <- summary(lm(Total.yield ~ consortium, df6.no))$coefficients[2, 1]

# Yield increase as an effect of treatment on microbiome (cwt/ac) is the product of the individual coefficients 
yield_inc <- append(yield_inc, (est_amd * est_cons))

# percent increase in total yield from treatment-related microbiome changes, relative to control  
pct <- est_amd * est_cons / mean(df6.no %>% filter(general_category == "Control") %>% pull(Total.yield)) * 100
pct_yield_inc <- append(pct_yield_inc, pct)
```


## Summary table /figure
```{r}
# add site, soil treatment, rotation length, and model type 
site <- c("OR", "ID", "ME1", "MI", "OR", "OR")
treatment <- c(rep("Amendment", times = 4), "Fumigation", "Mustard")
rotation <- c("2-yr", "3-yr", "2-yr", "3-yr", "3-yr", "3-yr")
modeltype <- c("lm", "SEM", "lm", "SEM", "lm", "SEM")

# bind these together with the model output data
model.table <- data.frame(cbind(site, treatment, rotation, modeltype, 
                                "p_value" = round(pval, 3), "mult_R2" = round(rsq, 2), 
                                "yield_inc" = round(yield_inc, 1), "pct_yield_inc" = round(pct_yield_inc, 1)))
# coerce back to numeric
for (i in 5:8){
  model.table[,i] <- as.numeric(model.table[,i])
}

# and here we are
model.table

write_tsv(model.table, file = "/Users/klas0061/Desktop/UMN/treatment_variance_modeling/figures_and_tables/supplemental_tables/microbiome_model_info_table.tsv")
```


## Conclusions

#### session info
```{r}
sessionInfo()
```